---
title: 'Monitoring forest regeneration - 1: processing ''2 poles coordinates'''
output:
  html_notebook: default
  html_document: default
---

# Introduction
2014 - Arborea - Pinus pinea plantation
4 (50 m * 50 m) plots
Each plot is divided in (10 m side) squared subplots
identified using the following scheme:

 "A5" "B5" "C5" "D5" "E5"  
 "A4" "B4" "C4" "D4" "E4"  
 "A3" "B3" "C3" "D3" "E3"  
 "A2" "B2" "C2" "D2" "E2"  
 "A1" "B1" "C1" "D1" "E1"  

In each subplot, point positions have been registered measuring the distances from the extremes of one side.
Let 'a' be the side used as reference for the positions measurements.
We will define a local cartesian sistem for each subplot using 'a' as 'x axis' and the left extreme of 'a', say 'A', as origin (0,0)
The coordinates of the point on opposite extreme of 'a', say 'B', are hence (10, 0)
For each registered point, say 'C', the distances (AC), say 'c', and (BC), say 'b', have been measured and registered.
To compute local cartesian coordinates of point 'C', say (lx, ly), one can exploit Erone's formula, actually reversing it.
(see: https://it.wikipedia.org/wiki/Formula_di_Erone 'Dimostrazione alternativa')
\[
\begin{eqnarray}
x & = & \frac {(a^{2}-b^{2}+c^{2})} {2a} \\
y & = & \frac {\sqrt{(4a^{2}c^{2}-(a^{2}-b^{2}+c^{2})^{2})}} {2a}
\end{eqnarray}
\]

## Access to data
```{r read_data}
library(tidyverse)
library(magrittr)
library(googlesheets)
library(spatstat)

suppressMessages(library(dplyr))
gs_ls()
tpp <- gs_title("DatiConcas")
gs_ws_ls(tpp)
ppwm <- tpp %>% gs_read(ws="RaccoltaFogliDati")
# Nei tabelloni originari (1 per plot!!) il calcolo delle coordinate cartesiane è sviluppato nei fogli 'cooridinate' che, in generale, pescano i valori da 'Dati'. I casi per i quali 'b'+'c'<10m sono però corretti direttamente in 'cooridinate'!! La tabella 'corr' raccolglie le differenze riscontrate tra 'Dati' e 'cooridinate'.
corr <- tpp %>% gs_read(ws="Correzioni",  col_types=cols(id_sogetto = col_character()))
# apply corrections
ppwm01 <- ppwm %>%
  left_join(corr) %>%
  mutate(b = ifelse(is.na(b_corr), b, b_corr),
         c = ifelse(is.na(c_corr), c, c_corr)) %>%
  select(-ends_with("_corr"))
```



# Compute plot-wise coordinates (x, y)
```{r process_coordinates}

# compute subplots origins
side <- 1:5
subplot_origin <- expand.grid(c=side, r=side) %>%
  mutate(subplot = paste(LETTERS[c], r, sep=""),
         x0 = 1000 * (c - 1),
         y0 = 1000 * (r - 1)) %>%
  select(-r, -c)

# distances are registered as cm (integers)
a = 10*100


ppwm01$`dbhNS`[is.na(ppwm01$`dbhNS`)==TRUE] <- ppwm01$`dbhEO`[is.na(ppwm01$`dbhNS`)==TRUE]
ppwm01$`dbhEO`[is.na(ppwm01$`dbhEO`)==TRUE] <- ppwm01$`dbhNS`[is.na(ppwm01$`dbhEO`)==TRUE]
ppwm01$`crown_rE`[is.na(ppwm01$`crown_rE`)==TRUE] <- 0
ppwm01$`crown_rO`[is.na(ppwm01$`crown_rO`)==TRUE] <- 0
ppwm01$`crown_rN`[is.na(ppwm01$`crown_rN`)==TRUE] <- 0
ppwm01$`crown_rS`[is.na(ppwm01$`crown_rS`)==TRUE] <- 0

ppwm02 <- ppwm01 %>%
# compute local cartesian coordiantes (lx, ly)
  mutate(lx = (a^2 - b^2 + c^2)/(2*a),
         ly = sqrt((4*a^2*c^2 - (a^2 - b^2 + c^2)^2))/(2*a),
         dbhMedio = sqrt(((dbhNS)^2+(dbhEO)^2)/2),
         area_crown = pi * (crown_rE+crown_rO+crown_rN+crown_rS)^2/16,
         area_fusto = pi * dbhMedio^2/4
         ) %>%
#(alternativa di Laura) ly = sqrt((2*(a^2*b^2 + a^2*c^2 + b^2*c^2)- (a^4 + b^4 + c^4))/(2*a)))
# compute plot-wise coordinates (x, y)
  full_join(subplot_origin) %>%
  mutate(x = (x0 + lx), 
         y = (y0 + ly))

# CREATE one PPP object for each 'plot'
cf <- 1/100 # conversion of "cm" to "m" 
#               marks = .[,7:16],
Arborea_permanent_plots <- ppwm02 %>%
  group_by(plot) %>%
#  filter(plot==1 & subplot=="A1") %$%
  do(ppp = ppp(.$x * cf, .$y * cf, 
               window = owin(xrange=c(round(min(.$x) * cf), round(max(.$x) * cf)),
                             yrange=c(0, round(max(.$y) * cf)  ), 
                               unitname="m"))
  ) %>%
  ungroup()
```


# Verify distribution of points
```{r plot}

plot(as.solist(Arborea_permanent_plots$ppp), main="2014Concas-Monitoring Pinus pinea regeneration, Arborea", main.panel=paste("Plot",Arborea_permanent_plots$plot))

plot.with.Axis <- function(i) {
  ungroup(Arborea_permanent_plots) %$% 
    plot(.$ppp[[i]], main=paste("Plot", .$plot[[i]]))
  at <- seq(0,50,10)
  Axis(side=1, at=at)
  Axis(side=4, at=at)
}
op <- par(no.readonly=T)
par(pin= c(par()$pin[1],.7 * par()$fin[2]))
for(i in 1:4) plot.with.Axis(i)
par(op)
  
#          dbh = ifelse(is.na(dbhEO), dbhNS, sqrt(mean(dbhNS^2, dbhEO^2)))

```


#Marks

```{r}

for (tid in 1:4) {
  
for (ty in c("h_tot", "age", "h_crown_ins", "dbhMedio", "area_crown", "area_fusto")) {
  
  trs <- ppp(ppwm02$x[ppwm02$plot==tid]*cf, ppwm02$y[ppwm02$plot==tid]*cf
           , xrange=c(round(min(ppwm02$x[ppwm02$plot==tid]) * cf), round(max(  ppwm02$x[ppwm02$plot==tid]) * cf)),
           yrange=c(0, round(max( ppwm02$y[ppwm02$plot==tid] ) * cf))
           , marks = ppwm02[ppwm02$plot==tid, ty])

  unitname(trs) <- "cm"
  plot(trs, arrange=FALSE, par(mar=rep(3,4)))

  if(ty == "h_tot") 
  mtext(paste(c("Plot", tid), collapse=" "), cex=1, outer=TRUE, line=-1)


  at <- seq(0,50,10)
  Axis(side=1, at=at)
  Axis(side=4, at=at)
  
}
  
  
  hist(ppwm02$age[ppwm02$plot==tid], main="age", xlab="age")
  hist(ppwm02$h_tot[ppwm02$plot==tid], main="h_tot", xlab="h_tot")
  hist(ppwm02$dbhMedio[ppwm02$plot==tid], main="dbhMedio", xlab="dbhMedio")

  
  }


```


