---
title: 'Monitoring forest regeneration - 1: processing ''2 poles coordinates'''
output:
  html_notebook: default
#  html_document: default
  pdf_document: default
---

# Introduction
2014 - Arborea - Pinus pinea plantation
4 (50 m * 50 m) plots
Each plot is divided in (10 m side) squared subplots
identified using the following scheme:

 "A5" "B5" "C5" "D5" "E5"  
 "A4" "B4" "C4" "D4" "E4"  
 "A3" "B3" "C3" "D3" "E3"  
 "A2" "B2" "C2" "D2" "E2"  
 "A1" "B1" "C1" "D1" "E1"  

In each subplot, point positions have been registered measuring the distances from the extremes of one side.
Let 'a' be the side used as reference for the positions measurements.
We will define a local cartesian sistem for each subplot using 'a' as 'x axis' and the left extreme of 'a', say 'A', as origin (0,0)
The coordinates of the point on opposite extreme of 'a', say 'B', are hence (10, 0)
For each registered point, say 'C', the distances (AC), say 'c', and (BC), say 'b', have been measured and registered.
To compute local cartesian coordinates of point 'C', say (lx, ly), one can exploit Erone's formula, actually reversing it.
(see: https://it.wikipedia.org/wiki/Formula_di_Erone 'Dimostrazione alternativa')
\[
\begin{eqnarray}
x & = & \frac {(a^{2}-b^{2}+c^{2})} {2a} \\
y & = & \frac {\sqrt{(4a^{2}c^{2}-(a^{2}-b^{2}+c^{2})^{2})}} {2a}
\end{eqnarray}
\]

## Access to data
```{r read_data}
library(tidyverse)
library(magrittr)
library(googlesheets)
library(spatstat)

suppressMessages(library(dplyr))
gs_ls()
tpp <- gs_title("DatiConcas")
gs_ws_ls(tpp)
plot_window <- tpp %>% gs_read(ws="Anagrafica")     
ppwm <- tpp %>% gs_read(ws="RaccoltaFogliDati")
# Nei tabelloni originari (1 per plot!!) il calcolo delle coordinate cartesiane è sviluppato nei fogli 'cooridinate' che, in generale, pescano i valori da 'Dati'. I casi per i quali 'b'+'c'<10m sono però corretti direttamente in 'cooridinate'!! La tabella 'corr' raccolglie le differenze riscontrate tra 'Dati' e 'cooridinate'.
corr <- tpp %>% gs_read(ws="Correzioni",  col_types=cols(id_sogetto = col_character()))

nais0 <- function(num) {
  return(ifelse(is.na(num), 0, num))
}
# apply corrections
ppwm01 <- ppwm %>%
  left_join(corr) %>%
  mutate(b = ifelse(is.na(b_corr), b, b_corr),
         c = ifelse(is.na(c_corr), c, c_corr)) %>%
  select(-ends_with("_corr"))
```
#define plots windows (old version)
plot_window <- tribble(
  ~plot, ~x00, ~y00, ~x01, ~y01,
      1,   0,   0,  5000,  5000,
      2,   0,   0,  5000,  5000,
      3,   0,   0,  5000,  4000,
      4,1000,   0,  2000,  5000
)


# Compute plot-wise coordinates (x, y)
```{r process_coordinates}

# compute subplots origins
side <- 1:5
subplot_origin <- expand.grid(c=side, r=side) %>%
  mutate(subplot = paste(LETTERS[c], r, sep=""),
         x0 = 1000 * (c - 1),
         y0 = 1000 * (r - 1)) %>%
  select(-r, -c)

# distances are registered as cm (integers)
a = 10*100


ppwm02 <- ppwm01 %>%
# compute local cartesian coordiantes (lx, ly)
  mutate(lx = (a^2 - b^2 + c^2)/(2*a),
         ly = sqrt((4*a^2*c^2 - (a^2 - b^2 + c^2)^2))/(2*a),
        #(alternativa di Laura) ly = sqrt((2*(a^2*b^2 + a^2*c^2 + b^2*c^2)- (a^4 + b^4 + c^4))/(2*a)))
         dbh = ifelse(is.na(dbhEO),dbhNS,sqrt((dbhNS^2+dbhEO^2)/2)),
         basal_area = pi * dbh^2/4,
         crown_area = pi * (nais0(crown_rE^2)+nais0(crown_rO^2)+nais0(crown_rN^2)+nais0(crown_rS^2))/4
         ) %>%
# compute plot-wise coordinates (x, y)
  full_join(subplot_origin) %>%
  mutate(x = (x0 + lx), 
         y = (y0 + ly))

# CREATE one PPP object for each 'plot'
cf <- 1/100 # conversion of "cm" to "m" 
#               marks = .[,7:16],

pw <- plot_window %>% 
  group_by(plot) %>% 
  do(window = owin(xrange=c(.$x00 * cf, .$x01 * cf), 
                   yrange=c(.$y00 * cf, .$y01 * cf),
                            unitname = 'm'))
Arborea_permanent_plots <- ppwm02 %>%
  full_join(plot_window) %>%
  group_by(plot) %>%
  do(ppp = ppp(.$x * cf, .$y * cf, 
               window = pw$window[.$plot][[1]],
               unitname="m",
               marks = data.frame(.$species, .$age, .$h_tot, 
                                  .$crown_area, .$dbh,
                                  .$basal_area, .$h_crown_ins))
  ) %>%
  ungroup()
```


# Verify distribution of points
```{r plot}
# x,y coorinates
plot(as.solist(Arborea_permanent_plots$ppp), main="2014Concas-Monitoring Pinus pinea regeneration, Arborea", main.panel=paste("Plot",Arborea_permanent_plots$plot), use.marks=F)

colnamesm = Arborea_permanent_plots$ppp[[1]] %>% marks() %>% colnames()
# marks
plot.with.Axis <- function(p, m) {
  ungroup(Arborea_permanent_plots) %$% 
  plot(.$ppp[[p]], main=paste("Plot", .$plot[[p]]), which.marks = m)
  mtext(colnamesm[m])
  at <- seq(0,50,10)
  Axis(side=1, at=at)
  Axis(side=4, at=at)
}
nm <- Arborea_permanent_plots$ppp[[1]] %>% marks() %>% colnames() %>% length()
op <- par(no.readonly=T)
par(pin= c(par()$pin[1],.7 * par()$fin[2]))
for(p in 1:4) for (m in 1:nm) plot.with.Axis(p, m)
par(op)
#  questa riga (nella versione di Davide del 30 ott) era sotto commento, non ho capitò perchè
dbh = ifelse(is.na(dbhEO), dbhNS, sqrt(mean(dbhNS^2, dbhEO^2)))

```


# Marks descriptive statistics

```{r descrpitive_statistics}
#rs calcolare statistiche individuali per plot (n.punti, n.NA, min. mean, max)
#    distribuzioni e correlazioni tra marche
#    DA FINIRE DI SISTEMARE

library(PerformanceAnalytics)

ow <- options("warn")[[1]]
options(warn = -1)
for(i in 1:4) chart.Correlation(marks(Arborea_permanent_plots$ppp[[i]])[, c(2:7)]
                                , histogram=TRUE, pch=19
                                , main=names(Arborea_permanent_plots$ppp)[i])

#  hist(ppwm02$age[ppwm02$plot==tid], main="age", xlab="age")
#  hist(ppwm02$h_tot[ppwm02$plot==tid], main="h_tot", xlab="h_tot")
#  hist(ppwm02$dbh[ppwm02$plot==tid], main="dbh", xlab="dbh")

  


```

# Statistical Analysis

```{r}

statsanalysis <- list()
for(tidplot in 1:4) {
  species = unique(toupper(ppwm02$species[ppwm02$plot==tidplot]))
  newstatsanalysis = tribble( ~species, ~mark,~NANumber, ~Media, ~Max, ~Min, ~CVPercentuale)
  for (spec in species) {
    for (c_mark in c("age", "h_tot", "crown_area", "dbh", "basal_area", "h_crown_ins")) {
    newstatsanalysis = rbind(newstatsanalysis, data.frame(
      species=spec,
      mark=c_mark,
      NANumber=sum(is.na(ppwm02[toupper(ppwm02$species)==spec & ppwm02$plot==tidplot, c_mark])),
      Media=mean(unlist(ppwm02[toupper(ppwm02$species)==spec & ppwm02$plot==tidplot, c_mark]), na.rm = T),
      Max=max(unlist(ppwm02[toupper(ppwm02$species)==spec & ppwm02$plot==tidplot, c_mark]), na.rm = T),
      Min=min(unlist(ppwm02[toupper(ppwm02$species)==spec & ppwm02$plot==tidplot, c_mark]), na.rm = T),
      CVPercentuale=sd(unlist(ppwm02[toupper(ppwm02$species)==spec & ppwm02$plot==tidplot, c_mark]), na.rm = T) /mean(unlist(ppwm02[toupper(ppwm02$species)==spec & ppwm02$plot==tidplot, c_mark]), na.rm = T)*100))
    }
  }
  statsanalysis[[tidplot]] <- newstatsanalysis
  show(paste("Plot",tidplot))
  show(statsanalysis[[tidplot]])
}

```


# Spatial Analysis

```{r}

for (tid in c(1,3)) {
  show(paste("Plot:", tid))
pltsm <- Arborea_permanent_plots$ppp[tid][[1]]

species = unique(pltsm$marks$..species)
for (s in species) {
  
  pltsm_s = pltsm[pltsm$marks$..species==s]
  
  marks(pltsm_s) <- pltsm_s$marks[3:4]

  show(paste("Specie:",s))
  
  mcf <- function(m1, m2) {m1*m2}
  try(plot(markcorr(pltsm_s, mcf)))
  try(plot(markcorr(pltsm_s)))
  try(plot(markvario(pltsm_s)))
  try(plot(markcrosscorr(pltsm_s)))

}
}

```
